---
title: "Final Project"
subtitle: "GEOG 351: Quantitative Analysis"
author: "Alice Clauss (260945446)"
format: revealjs
editor: visual
bibliography: references.bib
---

## Stages

-   Background

-   Methods

-   Results

-   Discussion

## Background: Data

My project focuses on the New Zealand Index of Multiple Deprivation (IMD), which was created in 2013 and updated in 2018, which is a ranking of the deprivation of by statistical area in New Zealand [@exeter2018]. I also used data from the 2018 New Zealand Census at the same geographic level.

```{r}
#| echo = FALSE

```

## Background: Index weighting

[![Figure 1. Weights of the IMD [@exeter2018].](IMD_2018_development.png){alt="Weights of the IMD" width="562"}](#0)

<https://quarto.org/docs/presentations/>.

## Bullets

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## Code

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Works cited
